---
title: "Spatial and Temporal Variability in Salamander Population Size"
author: "Allison K. Williams & Daniel J. Hocking"
output: word_document
---

<!--
output: powerpoint_presentation
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.width = 10, fig.height = 6)
```

```{r}
# Load packages
library(dplyr)
library(ggplot2)
library(readr)
library(lubridate)
library(here)
library(tidyr)
library(oSCR)
library(kableExtra)

# load all
load(here::here("analysis", "results", "mod_results.RData"))
```

## Objectives

- Estimate red-backed salamander density
- Estimate average home range size
- Compare spatial and temporal variation in density
- Examine the effects of rain and temperature on activity

change these as you see fit

## Background

- Bullet 1
- Bullet 2
- Bullet 3

## Background

More background

## Location

Western MD, forest type, blah blah blah

## Study Design

- 6 plots
- 50 boards/plot on 1 m grid
- SCR with VIE
- Spring & Fall 2017-2019 (6 sessions)
- 3-7 surveys per session

## Methods

More methods

## Methods - model

* Density estimated independently by plot-period (=36 sessions)
* Capture prob. ~ behavior + soil temp + precip + temp x precip

## Results

```{r}
kable(print(m8))
```

## Results

We had a total of `r nrow(pcin_1)` captures of `r length(unique(pcin_1$id))` individuals across all plots and surveys. 

```{r}
pcin_1_oscr$scrFrame
```

```{r, eval = FALSE}
pcin_1 %>%
  group_by(id) %>%
  dplyr::select(id) %>%
  summarise(count = n())
```

## Results

Sex Ratios - full csv saved to results

```{r}
# Sex ratios
sex_samps <- sapply(pcin_1_oscr$scrFrame$indCovs, function(x) table(x$sex))
sex_samps <- rbind(sex_samps, round(sex_samps[1,] / apply(sex_samps,2,sum),2))
dimnames(sex_samps) <- list(c("F","M","Prop.F"), paste0("S",1:36))
kable(as.data.frame(sex_samps[ , 1:6]))
write.csv(sex_samps, file = here::here("analysis", "results", "sex_ratios.csv"))
```

## Results

Density estimates per plot over the six sessions

```{r}
topmod <- m8
if(testing) {
  d_factor = 1
} else {
  d_factor = 4
}

#make a dataframe of values for DENSITY predictions
d.pred.df <- data.frame(session = rep(factor(1:length(tdf)), 1)) #session specific
#now predict
d_preds <- get.real(model = topmod, type = "dens", newdata = d.pred.df, d.factor = d_factor)
d_preds$sex <- factor(d_preds$sex)
levels(d_preds$sex) <- c("Female","Male")
# head(d_preds)

session_df <- session_df %>%
  dplyr::mutate(Session = session,
                session = as.integer(as.factor(Session)))

d_preds <- d_preds %>%
  dplyr::mutate(session = as.integer(as.factor(session))) %>%
  left_join(session_df)

density_summary_session <- d_preds %>%
  dplyr::select(date, pp, plot, estimate) %>%
  distinct() %>%
  group_by(date, plot) %>%
  dplyr::summarise(pp = mean(pp),
                   adult_density = sum(estimate))

density_summary <- density_summary_session %>%
  ungroup() %>%
  dplyr::select(adult_density) %>%
  dplyr::summarise(mean = mean(adult_density))

```

The mean density is over time and space is `r round(mean(density_summary_session$adult_density), 2)`$\pm$ `r round(sd(density_summary_session$adult_density), 2)`.

This means that we would expect `r round(5350 * mean(density_summary_session$adult_density), 0)` red-backed salamanders in an area of this forest the size of a football field.

```{r}
# add facet grid to separate plots and primary periods
g1 <- ggplot(d_preds, aes(x=pp, y=estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=4, position = position_dodge(width=0.5)) +
  facet_wrap(~plot) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Session") + ylab(expression(Density~(per~m^2)))

ggsave("analysis/figures/density_session.pdf", plot = g1, width = 8, height = 5)


g2 <- ggplot(d_preds, aes(x=date, y=estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=40)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=40)) +
  geom_point(size=4, position = position_dodge(width=40)) +
  facet_wrap(~plot) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Date") + ylab(expression(Density~(per~m^2))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("analysis/figures/density_date.pdf", plot = g2, width = 8, height = 5)

g2
```

## Spatial and Temporal Variation

```{r}
# Spatial and Temporal Variance/SD (in the means)

# variation among plots
plot_means <- d_preds %>%
  group_by(plot) %>%
  summarise(mean = mean(estimate, na.rm = TRUE),
            sd = sd(se, na.rm = TRUE))

spatial_sd <- sd(plot_means$mean)

# variation among sessions (primary periods)
session_means <- d_preds %>%
  group_by(session) %>%
  summarise(mean = mean(estimate, na.rm = TRUE),
            sd = sd(se, na.rm = TRUE))

temporal_sd <- sd(session_means$mean) # variation over time (averaged over space)
```

- Variation (SD) in density among plots averaged over time: `r spatial_sd`
- Variation (SD) in density among sessions averaged over plots: `r temporal_sd`

## Results

Movement and Home Range (~90% kernel density)

```{r}
s_pred_df <- data.frame(sex = rep(c("female", "male"), each = 1))

#now predict
s_preds <- get.real(model = topmod, type = "sig" , newdata = s_pred_df)
s_preds$sex <- factor(s_preds$sex)
levels(s_preds$sex) <- c("Female","Male")
head(s_preds)

# calculate 90% Home Range - delta method for uncertainty
s_preds <- s_preds %>%
  dplyr::mutate(hr90_estimate = pi * (estimate * 1.282) ^ 2,
                hr90_sd = sqrt(se^2 * (2 * pi * estimate)^2),
                hr90_lwr = hr90_estimate - 1.96 * hr90_sd,
                hr90_upr = hr90_estimate + 1.96 * hr90_sd)

g3 <- ggplot(s_preds, aes(x=sex, y=estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=4, position = position_dodge(width=0.5)) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Session") + ylab(expression("Sigma"~(m)))

ggsave("analysis/figures/sigma_session.pdf", plot = g3, width = 8, height = 5)

g3

g4 <- ggplot(s_preds, aes(x=sex, y=hr90_estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymin=hr90_lwr,ymax=hr90_upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=4, position = position_dodge(width=0.5)) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Session") + ylab(expression(Home~Range~(m^2)))

ggsave("analysis/figures/home_range_session.pdf", plot = g4, width = 8, height = 5)

# g4

# g5 <- ggplot(data.frame(x = c(4, 6)), aes(x = x)) +
#  stat_function(fun = dnorm, n = 101, args = list(mean = s_preds$hr90_estimate[1], sd = s_preds$hr90_sd[1]), geom = "area", fill = "red", alpha = 0.5) +
#   stat_function(fun = dnorm, n = 101, args = list(mean = s_preds$hr90_estimate[2], sd = s_preds$hr90_sd[2]), geom = "area", fill = "blue", alpha = 0.5) +
#   theme_bw() +
#   guide_legend(title = "Sex",
#                ) # scale_color_fivethirtyeight() +
#   xlab("Session") + ylab(expression(Home~Range~(m^2)))
  
# library(MASS)       # for fitdistr(...)
hr90 <- data.frame(sex = s_preds$sex, hr90 = rnorm(100000, mean = s_preds$hr90_estimate, sd = s_preds$hr90_sd)) %>%
  arrange(sex, hr90)

dat <- with(density(hr90[which(hr90$sex == "Female"), "hr90"], kernel = c("gaussian")), data.frame(x, y)) %>%
  dplyr::rename(hr90 = x, freq = y) %>%
  dplyr::mutate(sex = "Female")

dat2 <- with(density(hr90[which(hr90$sex == "Male"), "hr90"], kernel = c("gaussian")), data.frame(x, y)) %>%
  dplyr::rename(hr90 = x, freq = y) %>%
  dplyr::mutate(sex = "Male")

dat <- bind_rows(dat, dat2)

g5 <- ggplot(data = dat, aes(x = hr90, y = freq, color = sex, group = sex)) +
  geom_area(aes(fill = sex), alpha = 0.3) +
  xlab(expression(Home~Range~(m^2))) +
  ylab("Kernel Density") + 
  theme_bw()

# g5 <- ggplot(data = dat, aes(x = hr90, y = ..density.., color = sex, group = sex)) +
#   stat_density(aes(fill = sex), alpha = 0.3) +
#   xlab(expression(Home~Range~(m^2))) +
#   ylab("Kernel Density") + 
#   theme_bw()

ggsave("analysis/figures/home_range_session.pdf", plot = g5, width = 8, height = 5)

g5
```


```{r, eval = FALSE}
s_pred_df <- data.frame(session = rep(factor(1:36),2),
                        sex = rep(c("female", "male"),each=36))
#now predict
s_preds <- get.real(model = topmod, type = "sig", newdata = s_pred_df)
s_preds$sex <- factor(s_preds$sex)
levels(s_preds$sex) <- c("Female","Male")
head(s_preds)

# calculate 90% Home Range - delta method for uncertainty
s_preds <- s_preds %>%
  dplyr::mutate(session = as.integer(as.factor(session)),
                hr90_estimate = pi * (estimate * 1.282) ^ 2,
                hr90_sd = sqrt(se^2 * (2 * pi * estimate)^2),
                hr90_lwr = hr90_estimate - 1.96 * hr90_sd,
                hr90_upr = hr90_estimate + 1.96 * hr90_sd) %>%
  left_join(session_df)


g6 <- ggplot(s_preds, aes(x=pp, y=estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymin=lwr,ymax=upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=4, position = position_dodge(width=0.5)) +
  facet_wrap(~plot) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Session") + ylab(expression(Sigma~(m)))

ggsave("analysis/figures/sigma_session.pdf", plot = g6)

g7 <- ggplot(s_preds, aes(x=pp, y=hr90_estimate, color = sex, group=sex)) +
  geom_line(size = 1, position = position_dodge(width=0.5)) +
  geom_errorbar(aes(ymin=hr90_lwr,ymax=hr90_upr), width=0, size=0.75, color=1,
                position = position_dodge(width=0.5)) +
  geom_point(size=4, position = position_dodge(width=0.5)) +
  facet_wrap(~plot) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Session") + ylab(expression(Home~Range~(m^2)))

ggsave("analysis/figures/home_range_session.pdf", plot = g7)

g7
```


## Results 

Soil temperature moderates the effect of rain. There is more surface activity with increasing rain at low temperatures but this effect goes away at higher temperatures and the trend may even reverse.

```{r}
#make a dataframe of values for DETECTION predictions
p_pred_df <- data.frame(rain = rep(seq(-1, 2.5, length.out = 100), 2),
                        soil = rep(c(-2, 2), each = 100),
                        b = 0) #, #pr(initial capture)
# sex=rep(c(0,1),each=100)) #binary sex
#now predict
p_preds <- get.real(model = topmod, type = "det", newdata = p_pred_df)
# p.preds$sex <- factor(p.preds$sex)
# levels(p.preds$sex) <- c("Female","Male")
p_preds$soil <- factor(p_preds$soil)
levels(p_preds$soil) <- c("Cold","Hot")

g8 <- ggplot(p_preds, aes(x = (rain * stds[which(stds$var == "rain"), "sd"]) + stds[which(stds$var == "rain"), "mean"], y = estimate, color = soil, group = soil)) +
 geom_ribbon(aes(ymin=lwr,ymax=upr, fill = soil), alpha=0.1) +
geom_line(size=1.5) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Rain 3-day total(cm)") + ylab("Initial Capture Prob.") + 
  coord_cartesian(xlim = c(0, 4))

ggsave("analysis/figures/detection_ain_soil.pdf", plot = g8, width = 8, height = 5)

g8
```

Another way to look at it with rain moderating the effect of temperature:

```{r}
#make a dataframe of values for DETECTION predictions
p_pred_df <- data.frame(soil = rep(seq(-2, 2, length.out = 100), 2),
                        rain = rep(c(-1, 2.5), each = 100),
                        b = 0) #, #pr(initial capture)
# sex=rep(c(0,1),each=100)) #binary sex
#now predict
p_preds <- get.real(model = topmod, type = "det", newdata = p_pred_df)
# p.preds$sex <- factor(p.preds$sex)
# levels(p.preds$sex) <- c("Female","Male")
p_preds$rain <- factor(p_preds$rain)
levels(p_preds$rain) <- c("Dry","Wet")

g9 <- ggplot(p_preds, aes(x = (soil * stds[which(stds$var == "soil"), "sd"]) + stds[which(stds$var == "soil"), "mean"], y = estimate, color = rain, group = rain)) +
 geom_ribbon(aes(ymin=lwr,ymax=upr, fill = rain), alpha=0.1) +
geom_line(size=1.5) +
  theme_bw() + # scale_color_fivethirtyeight() +
  xlab("Soil Temperature (C)") + ylab("Initial Capture Prob.")

ggsave("analysis/figures/detection_soil_rain.pdf", plot = g9, width = 8, height = 5)

g9
```

## Discussion

Compare to other papers and interpretation

## Future Directions

- Robust Design:
  - Temporary Emigration
  - Survival (~f(seasonal conditions))
- Bayesian Implementation
- Reproductive effort over time
- Growth rates
- Interactions with other species
- Comparison across range of red-backed salamander (SPARCnet)

## Thank You

- Acknowledgments (Jacey Brooks, Natalie Haydt, Elizabeth Myers, Jason Shaffer, Erin Gaylord, Justin Hansen, Jess Raney, Brady Flowers, Laura Smith, Nina Grove, Kenny Weaver, Shelbi Sullivan, Frostburg State University Biology Department and Students)
- This can be done with bullets or just pictures - don't need everyone

